\documentclass[11pt,letterpaper]{article}

\usepackage{amsmath}
\usepackage{multirow}
\usepackage[shortlabels]{enumitem}

\begin{document}

\title{CS 542 (Fall 2023) Written Assignment 1\\Bayes' Theorem and Na\"ive Bayes Classification}
\author{Due September 7, 2023}
\date{}
\maketitle

\section{Bayes' Theorem}

You get an email.  You know that 90\% of your email is legitimate ($L$) while 10\% is spam ($S$).

\begin{enumerate}[a.]
	\item Assume the following probabilities:
	\begin{itemize}
		\item The probability that an email contains the word ``Bitcoin'' ($B$) if it is spam is 96\%.
		\item The probability that an email contains the word ``Bitcoin'' if it is legitimate is 5\%.
	\end{itemize}
	What is the probability that the your new email is spam given that it contains the word ``Bitcoin''? \textbf{Show your work!}
	\item Assume the following probabilities:
	\begin{itemize}
		\item The probability that an email contains the word ``Covid'' ($C$) if it is spam is 50\%.
		\item The probability that an email contains the word ``Covid'' if it is legitimate is 12\%.
	\end{itemize}
	What is the probability that the your new email is legitimate given that it contains the word ``Covid''? \textbf{Show your work!}
\end{enumerate}

\section{Na\"ive Bayes}

The following problem is from the Jurafsky and Martin book, Exercise 4.2, reproduced below.\\

\noindent Given the following short movie reviews, each labeled with a genre, either comedy or action:

\begin{center}
\begin{tabular}{|c|c|}
\hline 
document & class \\ 
\hline 
fly fast shoot love & action \\ 
\hline 
fun couple love love & comedy \\ 
\hline 
fast furious shoot & action \\ 
\hline 
couple fly fast fun fun & comedy \\ 
\hline 
furious shoot shoot fun & action \\ 
\hline 
\end{tabular} 
\end{center}

\noindent and a new document D:\\

fast couple shoot fly\\

\noindent compute the most likely class for D. Assume a naive Bayes classifier and use add-1 smoothing for the likelihoods.\\

\noindent \textbf{Show your work!} In particular, show all of the probability distributions involved in the model (namely, $P(\mathrm{class})$ and $P(\mathrm{feature|class})$) and all of the steps used to calculate them. Create (conditional) probability tables such as those shown below.

\begin{center}
\begin{tabular}{|c|c|}
\hline 
class & $P(\mathrm{class})$ \\ 
\hline 
action &  \\ 
\hline 
comedy &  \\ 
\hline 
\end{tabular} 
\end{center}

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline 
\multicolumn{2}{|c|}{\multirow{2}{*}{$P(\mathrm{feature|class})$}} & \multicolumn{4}{|c|}{feature} \\ 
\cline{3-6}
\multicolumn{2}{|c|}{} & fast & couple & shoot & fly \\ 
\hline 
\multirow{2}{*}{class} & action &  &  &  & \\ 
\cline{2-6}
& comedy &  &  &  & \\ 
\hline
\end{tabular} 
\end{center}

Perform Laplace Smoothing to account for words that do not appear in one class.

\section*{Submission Instructions}

Please submit your solutions (in PDF format - printed and scanned images are OK) to the drop box on Canvas.

\end{document}